Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2015-01-20 algorithm.tex
AA: errori segnalati da reviewer 2 corretti. Rilettura ancora da fare

\item \label{pers02.enu:novelty_score_of_a_configuration}
The test configurations that are dominated, i.e. do not belong to $\mathscr{P}_i$, are neglected, while an \emph{innovation score} is given to the other ones, computed on the base of their separation from the non-dominated configurations of the past eras. Recalling that $s$ is the separation (see \defR{separation}), the innovation score of a non-dominated test configuration $\mathbf{c}^*$ is defined as
	\[
	is\left(\mathbf{c}^*\right)=\min\left\{ \left.s\left(\mathbf{c}\rightarrow\mathbf{c} \right)\right|\mathbf{c}\in\mathscr{P}_{i-1}\right\} 
	\]
In other words, a configuration $\mathbf{c}^*$ is characterized by as much innovation
as more separated it is from the configurations of the previous era.

@@

\item The goal of this step is preparing the regions that will compose the configuration space partition of the next era, i.e. $\mathscr{R}_{i+1}$. The following operations are performed:
	\begin{enumerate}
		\item Each high innovation region $R\in\mathscr{R}_{i,h}$ is split to obtain $\psi(R)$, as stated in \defR{splitting-a-region}. We use $\psi\left(\mathscr{R}_{i,h}\right)$ to denote the set of regions obtained as explained.
		\item Low innovation regions are left unchanged.
		\item No innovation regions are merged, two by two. To do so, pairs of no innovation contiguous regions $(R_1,R_2)$ are selected. We use $\mathscr{M}_i$ to represent the set of these pairs. This set is constructed such that pairs are independent, i.e. there cannot be two of them sharing a region. In other words, taking whatever $(R_1,R_2),(R_3,R_4) \in \mathscr{M}_i$, $R_i \neq R_j, \forall i,j\in\lbrace1,2,3,4\rbrace, i\neq j$. 
We merge each pair of these regions, i.e. $\forall (R_1,R_2)\in\mathscr{M}_i$ we calculate $\mu(R_1,R_2)$, according to \defR{merging-regions}.
We stress that the regions that belong to each selected pair must be contiguous, otherwise it is impossible to merge them. We use $\mu(\mathscr{M}_i)$ to denote the set of the merged regions, i.e. $\mu(\mathscr{M}_i) = \bigcup_{(R_1,R_2)\in\mathscr{M}_i} \mu(R_1,R_2)$
	\end{enumerate}

@@

On the other side, parameter $\alpha$ determines when a region is considered of high
innovation. High values of $\alpha$ mean that a region will hardly be
elected as high-innovation, yielding less splitting operations. This
slows down zooming in on portions of configuration space and make the
distribution of the simulated configuration more uniform, as there are
fewer small regions with a high test configuration concentration. The
opposite happens when small values of $\alpha$ are used. In this
introductory work we are not explicitly focused on investigating how
these parameters could be fine tuned, nevertheless
we conducted some preliminary tests to confirm the relative stability
of $PS$ with different choices and to select some ``reasonable'' average
values to perform the design space exploration, as it will discussed
in~\secR{ee}.

@@

Opposite to the ``Island  scheme'', a novel parallelization technique
permitted by PS can consist of different instances of PS, each one
running on the whole configuration space but having its own
configuration space partition that separately evolves. Each PS
instance independently performs the zoom-out and zoom-in in on
configuration space, splitting and merging its own regions. All PS
instances share the indication of the configurations already tested,
so that a configuration already tested by some PS instance, will not
be tested again by some other PS instance. In addition, PS instances
share the set of non-dominated configuration. Therefore, when a PS
instance evaluates the innovation score of its regions (steps
\ref{pers02.enu:novelty_score_of_a_configuration} and
\ref{enu:region_innovation_score} of \secR{steps}), it considers the
solutions found by the other PS instances. In this way, the
configuration space is spanned in parallel by multiple PS instances,
but, although autonomous, they are not isolated; on the contrary, they
continuously exchange information and influence each other, thanks to
sharing the already simulated configurations and the non-dominated
one. Otherwise stated, all the PS instances (independently but
jointly) contribute at the formation of the Pareto-set.

>>>>

\item \label{pers02.enu:novelty_score_of_a_configuration}
The test configurations that are dominated, i.e. do not belong to $\mathscr{P}_i$, are neglected, while an \emph{innovation score} is given to the other ones, computed on the base of their separation from the non-dominated configurations of the past eras. Recalling that $s$ is the separation (see \defR{separation}), the innovation score of a non-dominated test configuration $\mathbf{c}^*$ is defined as
	\[
	is\left(\mathbf{c}^*\right)=\min\left\{ \left.s\left(\mathbf{c}^*\rightarrow\mathbf{c} \right)\right|\mathbf{c}\in\mathscr{P}_{i-1}\right\} 
	\]
In other words, a configuration $\mathbf{c}^*$ is characterized by as much innovation
as more separated it is from the configurations of the previous era.

@@

\item The goal of this step is preparing the regions that will compose the configuration space partition of the next era, i.e. $\mathscr{R}_{i+1}$. The following operations are performed:
	\begin{enumerate}
		\item Each high innovation region $R\in\mathscr{R}_{i,h}$ is split to obtain $\psi(R)$, as stated in \defR{splitting-regions}. We use $\psi\left(\mathscr{R}_{i,h}\right)$ to denote the set of regions obtained as explained.
		\item Low innovation regions are left unchanged.
		\item No innovation regions are merged, two by two. To do so, pairs of no innovation contiguous regions $(R_1,R_2)$ are selected. We use $\mathscr{M}_i$ to represent the set of these pairs. This set is constructed such that pairs are independent, i.e. there cannot be two of them sharing a region. In other words, taking whatever $(R_1,R_2),(R_3,R_4) \in \mathscr{M}_i$, $R_i \neq R_j, \forall i,j\in\lbrace1,2,3,4\rbrace, i\neq j$. 
We merge each pair of these regions, i.e. $\forall (R_1,R_2)\in\mathscr{M}_i$ we calculate $\mu(R_1,R_2)$, according to \defR{merging-regions}.
We stress that the regions that belong to each selected pair must be contiguous, otherwise it is impossible to merge them. We use $\mu(\mathscr{M}_i)$ to denote the set of the merged regions, i.e. $\mu(\mathscr{M}_i) = \bigcup_{(R_1,R_2)\in\mathscr{M}_i} \mu(R_1,R_2)$
	\end{enumerate}

@@

On the other side, parameter $\alpha$ determines when a region is considered of high
innovation. High values of $\alpha$ mean that a region will hardly be
elected as high-innovation, yielding less splitting operations. This
slows down zooming in on portions of configuration space and makes the
distribution of the simulated configuration more uniform, as there are
fewer small regions with a high test configuration concentration. The
opposite happens when small values of $\alpha$ are used. In this
introductory work we are not explicitly focused on investigating how
these parameters could be fine tuned, nevertheless
we conducted some preliminary tests to confirm the relative stability
of $PS$ with different choices and to select some ``reasonable'' average
values to perform the design space exploration, as we will discuss
in~\secR{ee}.

@@

Opposite to the ``Island  scheme'', a novel parallelization technique
permitted by PS can consist of different instances of PS, each one
running on the whole configuration space but having its own
configuration space partition that separately evolves. Each PS
instance independently performs the zoom-out and zoom-in on
configuration space, splitting and merging its own regions. All PS
instances share the indication of the configurations already tested,
so that a configuration already tested by some PS instance, will not
be tested again by some other PS instance. In addition, PS instances
share the set of non-dominated configuration. Therefore, when a PS
instance evaluates the innovation score of its regions (steps
\ref{pers02.enu:novelty_score_of_a_configuration} and
\ref{enu:region_innovation_score} of \secR{steps}), it considers the
solutions found by the other PS instances. In this way, the
configuration space is spanned in parallel by multiple PS instances,
but, although autonomous, they are not isolated; on the contrary, they
continuously exchange information and influence each other, thanks to
sharing the already simulated configurations and the non-dominated
one. Otherwise stated, all the PS instances (independently but
jointly) contribute at the formation of the Pareto-set.

### 2015-01-20 concepts.tex
AA: errori segnalati da reviewer 2 corretti. Rilettura ancora da fare

In this section and the following one, we will deliberately keep our formulation as general as possible, even if the focus of this paper is the embedded system design space exploration. The goal is to make the generality of PS emerge. We argue that it is a strong point for PS, as for genetic algorithms, since (i) it permits to apply it to a wide range of different problems and (ii) it can be effectually used when the embedded system design is particularly challenging and no much a-priori knowledge is available to the designer. As already pointed in \secL{related_work}, the generality of genetic algorithms has made them widely deployed with respect to the other approaches.
Thanks to the generality of PS, we will not need to modify or adapt it in order to apply it to the case of embedded system design, as we will show in \secL{ee}. On the one hand, this legitimate our choice of keeping formulation general. On the other hand, more importantly, this confirms the capability of PS to automatically adapt to the problem under investigation, without any explicit intervention or knowledge of the designer, required, on the contrary, by~\cite{givargis_tvlsi02,santosh_hptr00,dellnitz2005covering} .

@@

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally, not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$, but only a subset 
is feasible. We call this subset {\em
feasible configuration space} of $S$ and indicate it as
$\mathcal{C}^*(S)$].

@@

When a dependency is present between parameters, it could be
misleading to use assumption about the ``meaning'' of
parameters in order to intuitively choose an appropriate value that
optimizes objectives.  It should be also pointed
out that it's often very difficult to identify dependencies using
analytical techniques and often designers have to use different
strategies to deal with such complexity, as seen
in~\secR{Related-work}. In this work we will often refer to
hardware/software dependencies, denoting dependencies occurring
between parameters belonging to the hardware and software categories,
as defined in~\secR{designspace}.

@@

In the design process, there are often a lot of ``local decisions'' in
which the designer can take advantage of well-known results or
heuristics. For example, if we consider a highly parallel
computer architecture in a configuration with very few registers, it is
well-known that increasing the number of ALUs does not bring tangible
better performances but only cause an increase in area occupation. An
intelligent exploration algorithm would not waste time evaluating
configurations with many ALUs and few registers. If $p_{1}$ is the
parameter ``number of registers'' and $p_{2}$ is the parameter
``number of ALUs'', we can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in P_{1}\times P_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is \emph{uninteresting}. This clearly shows that there are cases when, if a parameter lies in
a certain range, there are ranges of the other one that are not worth being explored. Intuitively, the cartesian product of these ranges is an \emph{uninteresting region}.

@@

We also consider, in measuring how much interesting a region is, the
\emph{innovation} that it adds. If, during a design space exploration,
a Pareto front approximation is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable innovation,
because the way it fulfills the objectives is similar to the other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is distant from previous ones is remarkable,
since it let the experimenter discover potential performance that have not yet considered before.
At each iteration, PS ``zooms in'' the most interesting regions while ``zooming out'' from the rest of configuration space. In order to implement this behavior algorithmically, at each iteration PS assigns the same number $N$ of simulations to each region. If a region $R$ is interesting, it is split in $k$ sub-regions before the next iteration starts. Therefore, in the next iteration, $N$ simulations are assigned to each sub-region. As a result, we are populating the original region $R$ with $k\cdot N$ simulations. On the contrary, uninteresting regions are merged, so that each one will receive less effort in the next iteration. In other words, ares of the configuration space that are more promising are split in small sub-regions, while the other ones are blended in big regions. As illustrated in \figR{small_and_big}, assigning $N$ simulations to each region makes small ones well populated while bigger ones sparsely explored.

@@

Since regions will be defined as cartesian products of parameter intervals, we need the following definition.
\begin{definition}[Parameter interval] Let $p_{i}$ be a parameter and $a_{i},b_{i}\in V_{i},a_{i}<b_{i}$.
A $p_{i}$- interval is
	\[
	\left[a_{i}..b_{i}\right]=\left[a_{i},b_{i}\right]\cap P_{i}
	\]
i.e. taking the interval $\left[a_{i}..b_{i}\right]$
is the set of values of $V_{i}$ lying in $\left[a_{i},b_{i}\right]$)
\end{definition}

@@

As already explained, we will split interesting regions and merge uninteresting ones. The region splitting and merging operations are defined starting from the corresponding operations defined on intervals.
\begin{definition}[Splitting an interval]
\defL{splitting-a-region}
Given a $p_i$ interval $\left[a_{i}..b_{i}\right]$ with more than one element, i.e. $\left|\left[a_{i}..b_{i}\right]\right|>1$, the splitting operation produces as result two contiguous intervals $\left[a_{i}..c_{i}\right],\left[d_{i}..b_{i}\right]$
 such that
	\begin{align}
		\begin{cases}
		\left[a_{i}..c_{i}\right]\cap\left[d_{i}..b_{i}\right] & =\emptyset\\
		\left[a_{i}..c_{i}\right]\cup\left[d_{i}..b_{i}\right] & =\left[a_{i}..b_{i}\right]\\
		\left|\left[a_{i}..c_{i}\right]\right| & =\left\lceil \frac{\left|\left[a_{i}..b_{i}\right]\right|}{2}\right\rceil 
		\end{cases}
	\end{align}

@@

\begin{definition}[Splitting a region]
\label{pers02.def:Splitting-a-region}
Consider a region $R=\prod\left[a_{i}..b_{i}\right]$.The split operator $\psi$ divides it into a set of sub-regions obtained splitting each $p_i$-interval and combining the resulting sub-intervals with the cartesian product, as shown in \figR{split}~(b). Formally:
	\begin{align}
		\psi(R) = \left\{ \left.\prod\left[x_{i}..y_{i}\right]\right|\left[x_{i}..y_{i}\right]\in \psi\left[a_{i}..b_{i}\right]  \right\} 
	\end{align}

@@

We now formally describe the merging operations.
\begin{definition}[Merging intervals]
Let $\left[a_{i}\dots b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$ be two contiguous $p_i$-intervals. The merge operation applied to them produces a new $p_i$-interval of the following form:
	\begin{align}
		\mu	 \left( \begin{array}{l}
				\left[a_{j}..y_{j}\right] \\
				\left[x_{j}..b_{j}\right]
		     \end{array} \right)
		=\begin{cases}
			\left[a_{j}..y_{j}\right] & \mbox{ if }b_{j}<x_{j}\\
			\left[x_{j}..b_{j}\right] & \mbox{ if }y_{j}<a_{j}
		\end{cases}
	\end{align}

>>>>

In this section and the following one, we will deliberately keep our formulation as general as possible, even if the focus of this paper is the embedded system design space exploration. The goal is to make the generality of PS emerge. We argue that it is a strong point for PS, as for genetic algorithms, since (i) it permits to apply it to a wide range of different problems and (ii) it can be effectually used when the embedded system design is particularly challenging and no much a-priori knowledge is available to the designer. As already pointed in \secR{Related-work}, the generality of genetic algorithms has made them widely deployed with respect to the other approaches.
Thanks to the generality of PS, we will not need to modify or adapt it in order to apply it to the case of embedded system design, as we will show in \secR{ee}. On the one hand, this legitimate our choice of keeping formulation general. On the other hand, more importantly, this confirms the capability of PS to automatically adapt to the problem under investigation, without any explicit intervention or knowledge of the designer, required, on the contrary, by~\cite{givargis_tvlsi02,santosh_hptr00,dellnitz2005covering} .

@@

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally, not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$, but only a subset 
is feasible. We call this subset {\em
feasible configuration space} of $S$ and indicate it as
$\mathcal{C}^*(S)$.

@@

When a dependency between parameters is present, it could be
misleading to use assumptions about the ``meaning'' of
parameters in order to intuitively choose an appropriate value that
optimizes the objectives.  It should be also pointed
out that it's often very difficult to identify dependencies using
analytical techniques and often designers have to use different
strategies to deal with such complexity, as seen
in~\secR{Related-work}. In this work we will often refer to
hardware/software dependencies, denoting dependencies occurring
between parameters belonging to the hardware and software categories,
as defined in~\secR{designspace}.

@@

In the design process, there are often a lot of ``local decisions'' in
which the designer can take advantage of well-known results or
heuristics. For example, if we consider a highly parallel
computer architecture in a configuration with very few registers, it is
well-known that increasing the number of ALUs does not bring tangible
better performances but only cause an increase in area occupation. An
intelligent exploration algorithm would not waste time evaluating
configurations with many ALUs and few registers. If $p_{1}$ is the
parameter ``number of registers'' and $p_{2}$ is the parameter
``number of ALUs'', we can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in V_{1}\times V_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is \emph{uninteresting}. This clearly shows that there are cases when, if a parameter lies in
a certain range, there are ranges of the other one that are not worth being explored. Intuitively, the cartesian product of these ranges is an \emph{uninteresting region}.

@@

We also consider, in measuring how much interesting a region is, the
\emph{innovation} that it adds. If, during a design space exploration,
a Pareto front approximation is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable innovation,
because the way it fulfills the objectives is similar to the other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is distant from previous ones is remarkable,
since it let the experimenter discover potential performance that have not yet considered before.
At each iteration, PS ``zooms in'' the most interesting regions while ``zooming out'' from the rest of configuration space. In order to implement this behavior algorithmically, at each iteration PS assigns the same number $N$ of simulations to each region. If a region $R$ is interesting, it is split in $k$ sub-regions before the next iteration starts. Therefore, in the next iteration, $N$ simulations are assigned to each sub-region. As a result, we are populating the original region $R$ with $k\cdot N$ simulations. On the contrary, uninteresting regions are merged, so that each one will receive less effort in the next iteration. In other words, areas of the configuration space that are more promising are split in small sub-regions, while the other ones are blended in big regions. As illustrated in \figR{small_and_big}, assigning $N$ simulations to each region makes small ones well populated while bigger ones sparsely explored.

@@

Since regions will be defined as cartesian products of parameter intervals, we need the following definition.
\begin{definition}[Parameter interval] Let $p_{i}$ be a parameter and $a_{i},b_{i}\in V_{i},a_{i}<b_{i}$.
A $p_{i}$- interval is
	\[
	\left[a_{i}..b_{i}\right]=\left[a_{i},b_{i}\right]\cap V_{i}
	\]
i.e. taking the interval $\left[a_{i}..b_{i}\right]$
is the set of values of $V_{i}$ lying in $\left[a_{i},b_{i}\right]$)
\end{definition}

@@

As already explained, we will split interesting regions and merge uninteresting ones. The region splitting and merging operations are defined starting from the corresponding operations defined on intervals.
\begin{definition}[Splitting an interval]
\defL{splitting-intervals}
Given a $p_i$ interval $\left[a_{i}..b_{i}\right]$ with more than one element, i.e. $\left|\left[a_{i}..b_{i}\right]\right|>1$, the splitting operation produces as result two contiguous intervals $\left[a_{i}..c_{i}\right],\left[d_{i}..b_{i}\right]$
 such that
	\begin{align}
		\begin{cases}
		\left[a_{i}..c_{i}\right]\cap\left[d_{i}..b_{i}\right] & =\emptyset\\
		\left[a_{i}..c_{i}\right]\cup\left[d_{i}..b_{i}\right] & =\left[a_{i}..b_{i}\right]\\
		\left|\left[a_{i}..c_{i}\right]\right| & =\left\lceil \frac{\left|\left[a_{i}..b_{i}\right]\right|}{2}\right\rceil 
		\end{cases}
	\end{align}

@@

\begin{definition}[Splitting a region]
\defL{splitting-regions}
Consider a region $R=\prod\left[a_{i}..b_{i}\right]$.The split operator $\psi$ divides it into a set of sub-regions obtained splitting each $p_i$-interval and combining the resulting sub-intervals with the cartesian product, as shown in \figR{split}~(b). Formally:
	\begin{align}
		\psi(R) = \left\{ \left.\prod\left[x_{i}..y_{i}\right]\right|\left[x_{i}..y_{i}\right]\in \psi\left[a_{i}..b_{i}\right]  \right\} 
	\end{align}

@@

We now formally describe the merging operations.
\begin{definition}[Merging intervals]
Let $\left[a_{i}\dots b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$ be two contiguous $p_i$-intervals. The merge operation applied to them produces a new $p_i$-interval of the following form:
	\begin{align}
		\mu	 \left( \begin{array}{l}
				\left[a_{i}..y_{i}\right] \\
				\left[x_{i}..b_{i}\right]
		     \end{array} \right)
		=\begin{cases}
			\left[a_{i}..y_{i}\right] & \mbox{ if }b_{i}<x_{i}\\
			\left[x_{i}..b_{i}\right] & \mbox{ if }y_{i}<a_{i}
		\end{cases}
	\end{align}

### 2015-01-20 introduction.tex
AA: errori segnalati da reviewer 2 corretti. Rilettura ancora da fare

A preliminary estimation of hardware/software dependencies in order to
reduce the design space is addressed in~\cite{Catania2008}, where
authors propose to deal with the software parameters separately,
statistically testing their average effect on objectives before the
actual exploration is performed. Also in~\cite{merging} authors
analyze the impact of including compilation parameters into the design
space exploration. The results of both works confirm the
important role that hardware/software dependencies play in a
multi-objective exploration and thus justify the choice of explicitly
including the two class of parameters in the design space without
making any distinction between them from the algorithm's perspective.
In fact, one of the purposes of the proposed PS algorithm is to avoid
the need of any a priori knowledge about those dependencies, letting
the algorithm itself to discover regions where interaction among
parameters generates new Pareto-optimal points (see the PS scoring
system described in the next sections).

Abraham, Rau and Schreiber proposed in \cite{santosh_hptr00} to
decompose the system under evaluation into components that interact
minimally with each other. Pareto sets for each component are found
and, provided that ``monotonicity'' exists, the complete Pareto set is
computed merging the component Pareto sets. Roughly speaking (see
section 4 of the same paper for more details), monotonicity property
guarantees that the best system can be obtained as a composition of
the best components. This approach would perfectly work if all the
components were perfectly isolated, i.e., if there were no
dependencies among them. But real scenarios rarely expose monotonicity
property thus resulting in inaccuracies, as stated in the same paper).

Other approaches, as \cite{fornaciari_codes01,palesi_iwsoc02}, are
based the concept of \emph{sensitivity analysis}, i.e. measuring of
how much the objective varies when varying each parameter.  Parameters
are ordered based on their sensitivity. The first parameter (the most
sensitive one) is varied, while keeping the other parameters fixed to
arbitrary values, and its best value is found. The limited accuracy
shown by these approaches can be explained by the limited and rigid
exploration of the parameter space: after fixing the best first
parameter value, there are no more chances to consider different
values. It is worth noticing that this approach can not capture
``local sensitivity'', i.e. the objectives may be more sensitive to
some ranges of a parameter and less sensitive to other ranges of the
same parameter. Moreover it can not capture ``combined sensitivity'',
i.e. the objectives may be more sensitive to a range of a parameter
$p_{1}$, only when other parameters are within certain ranges, and
less sensitive to the same range of $p_{1}$ when the other parameters
are in different ranges. Otherwise, our algorithm can both capture
local and combined sensitivity.

>>>>

A preliminary estimation of hardware/software dependencies in order to
reduce the design space is addressed in~\cite{Catania2008}, where
authors propose to deal with the software parameters separately,
statistically testing their average effect on objectives before the
actual exploration is performed. Also in~\cite{merging} authors
analyze the impact of including compilation parameters into the design
space exploration. The results of both works confirm the
important role that hardware/software dependencies play in a
multi-objective exploration and thus justify the choice of explicitly
including the two classes of parameters in the design space without
making any distinction between them from the algorithm's perspective.
In fact, one of the purposes of the proposed PS algorithm is to avoid
the need of any a priori knowledge about those dependencies, letting
the algorithm itself to discover regions where interaction among
parameters generates new Pareto-optimal points (see the PS scoring
system described in the next sections).

Abraham, Rau and Schreiber proposed in \cite{santosh_hptr00} to
decompose the system under evaluation into components that interact
minimally with each other. Pareto sets for each component are found
and, provided that ``monotonicity'' exists, the complete Pareto set is
computed merging the component Pareto sets. Roughly speaking (see
section 4 of the same paper for more details), monotonicity property
guarantees that the best system can be obtained as a composition of
the best components. This approach would perfectly work if all the
components were perfectly isolated, i.e., if there were no
dependencies among them. But real scenarios rarely expose monotonicity
property thus resulting in inaccuracies, as stated in the same paper.

Other approaches, as \cite{fornaciari_codes01,palesi_iwsoc02}, are
based on the concept of \emph{sensitivity analysis}, i.e. measuring of
how much the objective varies when varying each parameter.  Parameters
are ordered based on their sensitivity. The first parameter (the most
sensitive one) is varied, while keeping the other parameters fixed to
arbitrary values, and its best value is found. The limited accuracy
shown by these approaches can be explained by the limited and rigid
exploration of the parameter space: after fixing the best first
parameter value, there are no more chances to consider different
values. It is worth noticing that this approach can not capture
``local sensitivity'', i.e. the objectives may be more sensitive to
some ranges of a parameter and less sensitive to other ranges of the
same parameter. Moreover it can not capture ``combined sensitivity'',
i.e. the objectives may be more sensitive to a range of a parameter
$p_{1}$, only when other parameters are within certain ranges, and
less sensitive to the same range of $p_{1}$ when the other parameters
are in different ranges. Otherwise, our algorithm can both capture
local and combined sensitivity.
